Naturalness in U(1) extensions of the MSSM

### 2015-01-23 paper/e6ssm-tuning-paper.tex
Added reference for gluino limits, updated table with new benchmark values, added short description of the constrained benchmark

\begin{table}[h]
\centering
\begin{ruledtabular}
\begin{tabular}{cccc}
& MSSM & E$_6$SSM BM1 & E$_6$SSM BM2 \\
\hline
$\tan\beta(M_Z)$ & 10 & 10 & 10 \\
$s(M_{\mathrm{SUSY}})$ [GeV] & $-$ & 6700 & 6700 \\
$\kappa_{1,2,3}(M_{\mathrm{SUSY}})$ & $-$ & 0.6 & 0.6 \\
$\lambda_{1,2}(M_{\mathrm{SUSY}})$ & $-$ & 0.2 & 0.2 \\
$\mu_{\mathrm{eff}}(M_{\mathrm{SUSY}})$ [GeV] & 689.7 & 1093.3 & -1093.3 \\
$B_{\mathrm{eff}}(M_{\mathrm{SUSY}})$ [GeV] & 345.7 & 3792.7 & -3792.7 \\
$A_t(M_{\mathrm{SUSY}})$ [GeV] & -3335.7 & -1100 & -3335.7 \\
$A_b(M_{\mathrm{SUSY}}), A_\tau (M_{\mathrm{SUSY}})$ [GeV] & 0 & 0 & 0 \\
$m_{Q3}^2(M_{\mathrm{SUSY}})$ [GeV$^2$] & $4.45379310\times 10^6$ & $4.49655\times 10^5$ & $6.5\times 10^5$ \\
$m_{u3}^2(M_{\mathrm{SUSY}})$ [GeV$^2$] & $4\times 10^6$ & $5.86207\times 10^5$ & $4.5\times 10^5$ \\
$M_1(M_{\mathrm{SUSY}})$ [GeV] & 300 & 300 & 300 \\
$M_2(M_{\mathrm{SUSY}})$ [GeV] & 2000 & 1050 & 2000 \\
$M_3(M_{\mathrm{SUSY}})$ [GeV] & 2000 & 2000 & 1200 \\
$M_1^\prime(M_{\mathrm{SUSY}})$ [GeV] & $-$ & 300 & 300 \\
\hline
$M_{Z^\prime}$ [GeV] & $-$ & 2473.2 & 2473.1 \\
$m_{h_1}$ [GeV] & 124.3 & 125.0 & 124.9 \\
$m_{\tilde{t}_1}$ [GeV] & 1942.1 & 993.8 & 839.6 \\
$m_{\tilde{t}_2}$ [GeV] & 2220.1 & 1174.8 & 1285.2 \\
$m_{\tilde{g}}$ [GeV] & 2259.8 & 2290.0 & 1453.8\\
\hline
$\Delta(M_X=20\textrm{ TeV})$ & 157.3 & 165.3 & 228.5 \\
$\Delta(M_X=10^{16}\textrm{ GeV})$ & 1089.0 & 1722.3 & 917.6
\end{tabular}
\end{ruledtabular}
\caption{Parameters and masses for the MSSM point and the E$_6$SSM benchmark points 
BM1 and BM2 for which the fine tuning as a function of cutoff scale is studied. For the
E$_6$SSM benchmarks we define $\mu_{\mathrm{eff}}\equiv \lambda s/\sqrt{2}$ and
$B_{\mathrm{eff}}=A_\lambda$. All other (diagonal) soft scalar squared masses not shown above are set to $2.5\times 10^7$ GeV$^2$, and all soft trilinears are set to zero. In the E$_6$SSM, we additionally set $\mu^\prime=5000$ GeV and $B\mu^\prime=5000$ GeV$^2$.}
\label{tab:benchmarks}
\end{table}

@@

By contrast the tuning for E$_6$SSM BM2 is very different.  This point
was chosen as it had much lighter gluino mass that is just
above the experimental limit of $1.4$ TeV \cite{}.  At $20$ TeV it was
not amongst the lowest tuned points, since at that scale the tree
level tuning from $M_{Z^\prime}$ dominates.  However the reduction in
$M_3$ means that $\Delta_{M_3}$ is substantially lower and only
becomes the dominant tuning very close to the GUT scale, giving a
tuning at $10^{16}$ GeV of $\approx 550$, which is far below that of the
other two benchmark points.

>>>>

\begin{table}[h]
\centering
\begin{ruledtabular}
\begin{tabular}{cccc}
& MSSM & E$_6$SSM BM1 & E$_6$SSM BM2 \\
\hline
$\tan\beta(M_Z)$ & 10 & 10 & 10 \\
$s(M_{\mathrm{SUSY}})$ [GeV] & $-$ & 6700 & 6700 \\
$\kappa_{1,2,3}(M_{\mathrm{SUSY}})$ & $-$ & 0.6 & 0.52 \\
$\lambda_{1,2}(M_{\mathrm{SUSY}})$ & $-$ & 0.2 & 0.13 \\
$\mu_{\mathrm{eff}}(M_{\mathrm{SUSY}})$ [GeV] & 689.7 & 1093.3 & 1313.0 \\
$B_{\mathrm{eff}}(M_{\mathrm{SUSY}})$ [GeV] & 345.7 & 3792.7 & 817.8 \\
$A_t(M_{\mathrm{SUSY}})$ [GeV] & -3335.7 & -1100 & -1103.2 \\
$m_{Q3}^2(M_{\mathrm{SUSY}})$ [GeV$^2$] & $4.45\times 10^6$ & $4.50\times 10^5$ & $3.61\times 10^6$ \\
$m_{u3}^2(M_{\mathrm{SUSY}})$ [GeV$^2$] & $4\times 10^6$ & $5.86\times 10^5$ & $2.04\times 10^6$ \\
$M_1(M_{\mathrm{SUSY}})$ [GeV] & 300 & 300 & 173.4 \\
$M_2(M_{\mathrm{SUSY}})$ [GeV] & 2000 & 1050 & 281.4 \\
$M_3(M_{\mathrm{SUSY}})$ [GeV] & 2000 & 2000 & 1200 \\
$M_1^\prime(M_{\mathrm{SUSY}})$ [GeV] & $-$ & 300 & 175.2 \\
\hline
$M_{Z^\prime}$ [GeV] & $-$ & 2473.2 & 2512.7 \\
$m_{h_1}$ [GeV] & 124.3 & 125.0 & 126.2 \\
$m_{\tilde{t}_1}$ [GeV] & 1942.1 & 993.8 & 1665.0 \\
$m_{\tilde{t}_2}$ [GeV] & 2220.1 & 1174.8 & 2094.4 \\
$m_{\tilde{g}}$ [GeV] & 2259.8 & 2290.0 & 1407.4\\
\hline
$\Delta(M_X=20\textrm{ TeV})$ & 157.3 & 165.3 & 402.1 \\
$\Delta(M_X=10^{16}\textrm{ GeV})$ & 1089.0 & 1722.3 & 546.7
\end{tabular}
\end{ruledtabular}
\caption{Parameters and masses for the MSSM point and the E$_6$SSM benchmark points 
BM1 and BM2. For the
E$_6$SSM we define $\mu_{\mathrm{eff}}\equiv \lambda s/\sqrt{2}$ and
$B_{\mathrm{eff}}=A_\lambda$. In BM1 all other (diagonal) soft scalar squared masses not shown above are set to $(5\mathrm{ TeV})^2$, the soft trilinears $A_b$, $A_\tau=0$, $\mu^\prime=5000$ GeV and $B\mu^\prime=5000$ GeV$^2$. In BM2 the parameters not shown above are set to the values obtained by running the cE$_6$SSM point with $m_0=2.2$ TeV, $A_0=500$ GeV, $M_{1/2}=1003$ GeV, $\kappa_{1,2,3}(M_X)=0.1923$, $\lambda(M_X)=0.2646$ to the SUSY scale, and setting $\mu^\prime(M_{\mathrm{SUSY}})=897.9$ GeV and $B\mu^\prime(M_{\mathrm{SUSY}})=-4.2\times10^5$ GeV$^2$.}
\label{tab:benchmarks}
\end{table}

@@

By contrast the tuning for E$_6$SSM BM2 is very different. This point
was chosen as it had much lighter gluino mass that is just
above the experimental limit of $1.4$ TeV \cite{Aad:2014lra}. 
At $20$ TeV this benchmark is not amongst the lowest tuned points, since at that scale the tree
level tuning from $M_{Z^\prime}$ dominates.  However the reduction in
$M_3$ means that $\Delta_{M_3}$ is substantially lower and only
becomes the dominant tuning at a much larger scale of $M_X\gtrsim 10^{12}-10^{13}$ GeV, giving a
tuning at $10^{16}$ GeV of $\approx 546$, which is far below that of the
other two benchmark points.

In addition to this, the soft parameters in BM2 follow a pattern similar to that found in the constrained
model. With the exception of the parameters $m_{Q3}^2$, 
$m_{u3}^2$, $m_{H_d}^2$, $m_{H_u}^2$, and $M_3$, the values of which are given in Table \ref{tab:benchmarks}, 
the soft masses at the SUSY scale correspond to the values that result in the cE$_6$SSM with $m_0=2.2$ TeV, 
$M_{1/2}=1003$ GeV, $A_0=500$ GeV, $\kappa_{1,2,3}(M_X)=0.1923$, $\lambda(M_X)=0.2646$ and
$\lambda_{1,2}(M_X)=0.1$. This leads to a significant reduction in the contributions to the RG running of $m_{H_u}^2$ and $m_{Q3}^2$ coming from terms of the form $g_1^2\Sigma_1$ and, to a lesser extent, $g_1'^2\Sigma_1'$. Here we define (see also Eq. (\ref{}))
\begin{align*}
\Sigma_1&=\sum_{i=1}^3\left ( m_{Qi}^2-2m_{ui}^2+m_{di}^2+m_{ei}^2-m_{Li}^2+m_{H_{ui}}^2-m_{H_{di}}^2+m_{\overline{D}_i}^2-m_{Di}^2\right )-m_{H'}^2+m_{\overline{H'}}^2,\\
\Sigma_1'&=\sum_{i=1}^3\left ( 6m_{Qi}^2+3m_{ui}^2+6m_{di}^2+m_{ei}^2+4m_{Li}^2-4m_{H_{ui}}^2-6m_{H_{di}}^2+5m_{Si}^2-9m_{\overline{D}_i}^2-6m_{Di}^2\right )\\
&\quad{}+4m_{H'}^2-4m_{\overline{H'}}^2.
\end{align*}
In the unconstrained case, this contribution acts to drive up the values of $m_{Q3}^2$ and $m_{H_u}^2$, and thus the associated tuning sensitivities, at the input scale $M_X$. In the case of BM2, on the other hand, the reduced splitting between the soft masses leads to a much smaller contribution from these terms. Together with the reduction in $M_3$ described above, this allows to maintain the observed low fine tuning at very large values of $M_X$.
